---
title: "Lasioglossum Of Puerto Rico And the Carebbean Species Distribution Model"
author: "Maylin Caldwell, Sofía Meléndez Cartagena"
date: "4/13/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
## What is this?
	Well this is the script for our BigData course! We plan on doing a species distribution model to determian the habitat suitability value for the genus Lasioglossum in Puerto Rico as well as adjecent Islands in the Caribbean.
## What should I now to get it done?
	I suggest reading and maybe working on these tutorials before getting started with the data:
	> https://rspatial.org/index.html
	> https://cengel.github.io/R-data-wrangling/
## An explanation on our required libraries [I will fill this out later]

# Get to coding

## Set WorkingDirectory
```{r}
# setwd("~/../TopicosFinalProyect") This will vary by computer
```


## Required Libraries
```{r}
library(tidyverse)
library(wallace)
```

## Document imports
```{r}
#This one looks pretty
dat <- read_csv("../Data/BeeData4Real.csv")
#This one is functional
dat2 <- read.csv("../Data/BeeData4Real.csv", stringsAsFactors = FALSE)

```

## Build funtions if needed
### Use one code block per function

	Function to Transform Hexadecimal Coordinates to Decimal
```{r}
Hexa2Deci <- function(degree,min,sec){
    degree + (min/60) + (sec/3600) 
}
```

## Variable calls
### It's a good practice to have your variables groupd together in blocks if they are meant to be used together. 
```{r}
latitud <- dat2$Latitud #Vector for manipulating latitud col
longitud <- dat2$Longitud # Vector for manipulatin longitud col
corX <- c("DX","MX","SX") # Holds Titles for spread latitud
cory <- c("DY","MY","SY") # Holds Titles for spread longitud
```

## Main Body of Code

### Data Clean up

	Use gsub to clean the special characters from latitud and longitud vectors. Save to LongitudClean and LatitudClean. Proper gsub usage goes as following
```{r}
# Don't run this
subVec <- gsub("Pattern","Substitute",Vector)
```

```{r}
NoSpecialDat <- dat %>%
    mutate(Latitud = LatitudClean, Longitud = LongitudClean)

str(NoSpecialDat)
```

	Spread the latitude and longitude columms in to columms that represent Degrees, Minutes, and Seconds
```{r}
SeperatedDat <- NoSpecialDat %>%
    separate(Latitud,corX,sep = "[.]") %>% 
    separate(Longitud,cory,sep = "[.]") 
str(SeperatedDat)
```

	Convert To interger
```{r}
IntDat <- SeperatedDat %>%
	mutate(DX = as.integer(DX), MX = as.integer(MX), SX = as.integer(SX)) %>%
	mutate(DY = as.integer(DY), MY = as.integer(MY), SY = as.integer(SY))
str(IntDat)
```

	NAs will probably show up here. We should be mindfull of the original data and use it to replace the missing values. The proper steps to verify this is
```{r}
#Don't run this
summary(dataframe)
holder <- is.na(dataframe)
dataframe[holder] #allows you to view all the NAs
which(dataframe[holder]) #tells you the posicion within the dataframe of the NAs decide what to do from here on out
```

	Connvert do Decadecimal
```{r}
latitud <- Hexa2Deci(IntDat$DX,IntDat$MX,IntDat$SX)
longitud <- Hexa2Deci(IntDat$DY,IntDat$MY,IntDat$SY)
```

	Make a DataFrame to hold your new coodinates based on our original dat. Then imput the new coordinates in to 	the data frame
```{r}
AbejaTrans <- dat
AbejaTrans$Latitud <- latitud 
AbejaTrans$Longitud <- -1*longitud
```

	Subset the data frame so we only keep the species and the coordinates.
```{r}
CleanDF <- AbejaTrans[,2:4]

CleanDF <- CleanDF %>%
  rename("name" = Species,"latitude" = Latitud,"longitude" = Longitud)
```

	Save data in to a CSV for future convinience
```{r}
write_csv(CleanDF, path = "../Output/abejaLimpia.csv")
```

	
	Now that out data is clean, let's run out species distribution app and follow it's workflow. Let's not forget export the code by the end of it all. 
```{r}
run_wallace()
```