---
title: "Lasioglossum Of Puerto Rico And the Carebbean Species Distribution Model"
author: "Maylin Caldwell, Sofía Meléndez Cartagena"
date: "4/13/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set WorkingDirectory
```{r}
# setwd("~/../TopicosFinalProyect") This will vary by computer
```


## Required Libraries
```{r}
library(tidyverse)
library(sdm)
library(raster)
library(rgdal)
```

## Document imports
```{r}
dat <- read_csv("../data/BeeData.csv", stringsAsFactors = FALSE)

```

## Build funtions if needed
### Use one code block per function

Function to Transform Hexadecimal Coordinates to Decimal
```{r}
Hexa2Deci <- function(degree,min,sec){
    degree + (min/60) + (sec/3600) 
}
```

## Variable calls
### It's a good practice to have your variables groupd together in blocks if they are meant to be used together. 
```{r}
latitud <- dat$Latitud
longitud <- dat$Longitud
corX <- c("DX","MX","SX")
cory <- c("DY","MY","SY")
```

## Main Body of Code

### Data Clean up

Use gsub to clean the special characters from latitud and longitud vectors. Save to LongitudClean and LatitudClean.

```{r}
NoSpecialDat <- dat %>%
    mutate(Latitud = LatitudClean, Longitud = LongitudClean) %>%
    str
```
Spread the latitude and longitude columms in to columms that represent Degrees, Minutes, and Seconds

```{r}
SeperatedDat <- NoSpecialDat %>%
    separate(Latitud,corX,sep = "[.]") %>% 
    separate(Longitud,cory,sep = "[.]") %>% 
    str
```
Convert To interger
```{r}
IntDat <- SeperatedDat %>%
	mutate(DX = as.integer(DX), MX = as.integer(MX), SX = as.integer(SX)) %>%
	mutate(DY = as.integer(DY), MY = as.integer(MY), SY = as.integer(SY)) %>%
	str 
```
Connvert do Decadecimal

```{r}
latitud <- Hexa2Deci(SeperatedDat$DX,SeperatedDat$MX,SeperatedDat$SX)
longitud <- Hexa2Deci(SeperatedDat$DY,SeperatedDat$MY,SeperatedDat$SY)
```